


```{r setup, echo=FALSE,results='hide', message=FALSE, warning=FALSE}

options(stringsAsFactors = FALSE)

# For the report
library(knitr)
library(mpmisc)
library(tbdbaid)

# For graphics
library(ggplot2)
library(gridExtra)

# For aggregation
library(plyr)
library(reshape2)


# Set default chunk options
opts_chunk$set(echo = FALSE,
               comment = "",
               message = FALSE,
               warning = TRUE,
               error = TRUE,
               fig.width=8,
               fig.height=5.6,
               dpi=300,
               out.width="700px",
               dev="png",
               cache = FALSE)



# Set some plot parameters that will want tweaking
point_size <- 3
line_size <- 1.3
colorwheel <- c("#31A354", "#ADDD8E", "#F03B20", "#43A2CA")



# Get dates for the latest quarter, and for its corresponding quarter two
# years back
first.qtr <- get_quarter(Date = Sys.Date(), offset = -9)
last.qtr <- get_quarter(Date = Sys.Date(), offset = -1)



# Set up a theme for all the plots
theme_tb <- theme_bw() +
            theme(legend.key.width = unit(.75, "inches"))




```


# Denver Metro TB Control Program
# Quarterly Strategic Area Review
# `r last.qtr$qtr`

Updated at `r paste(Sys.time())`

## TB Cases in the Denver Metropolitan Area
-------------------------------------------

```{r cases}

# Get active cases over the last ten years.
# Ensure that the oldest year is a full year
actives <- query_actives(start_date = as.Date(paste(as.numeric(format(Sys.Date(), "%Y")) - 10,
                                                    "-01-01", 
                                                    sep = "")),
                         stop_date = last.qtr$end.date)

metroactives <- subset(actives, metro_case %in% 'True')


# Create a quarter label
metroactives$plot_qtr <- with(metroactives, paste(yr_id, " Q", qtr_id, sep = ""))

# Create plot labels for county
metroactives$plot_group <- NA
metroactives$plot_group[metroactives$report_county %in% "Denver"] <- "Denver"
metroactives$plot_group[!metroactives$report_county %in% "Denver"] <- "Other Metro Counties"


# Aggregate into two years of quarterly counts and ten years of annual counts
actives_qtr <- count(metroactives[metroactives$date_id > first.qtr$start.date, ], "plot_qtr")

actives_annual <- count(metroactives, "yr_id")




```


```{r case_rates}

# Read in the population estimates
pop <- read.csv("data/metro_pop_estimates.csv")

# Aggregate
metro_pop <- ddply(pop, .var = "year", .fun = summarise, metro_pop = sum(population))


actives_rate <- merge(x = actives_annual,
                      y = metro_pop,
                      by.x = "yr_id",
                      by.y = "year",
                      all.x = TRUE)

# Fill in any years with missing population estimates using the previous years
actives_rate$metro_pop[is.na(actives_rate$metro_pop)] <-
  actives_rate$metro_pop[!is.na(actives_rate$metro_pop)][sum(!is.na(actives_rate$metro_pop))]


# Calculate rates
actives_rate$rate_pht <- with(actives_rate, freq / metro_pop * 1e5)

# For the most recent year, annualize the rate
actives_rate$rate_pht[actives_rate$yr_id %in% substr(last.qtr$qtr, 1, 4)] <- 
    with(actives_rate[actives_rate$yr_id %in% substr(last.qtr$qtr, 1, 4), ],
         freq / metro_pop * (4 / last.qtr$qtr.num * 1e5)
)


# Plot it
# ggplot(actives_rate, aes(x = plot_qtr, y = rate_pht_annualized, group = 1)) +
#   geom_point(size = point_size, color = "#31A354") +
#   geom_line(size = line_size, color = "#31A354") +
#   expand_limits(y = 0) +
#   labs(x = "Quarter identified (earliest of date treatment started or date case reported)", 
#        y = "Annualized rate per 100,000 residents") +
#   theme_tb
```


### TB Cases by County of Residence


```{r cases_by_county}

ggplot(metroactives, aes(x = yr_id, fill = plot_group)) +
    geom_bar(color = "black") +
    stat_bin(geom="text", aes(label=..count.., vjust = 1.25)) +
    scale_fill_manual("Patient's Residence", values = c("#ADDD8E", "#31A354")) +
    labs(x = "Year Identified (earliest of date treatment started or date case reported)", 
         y = "Number of cases") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))

```



```{r cases_table,results='asis'}


actives_print <- ddply(metroactives, .var = "yr_id", .fun = summarise,
                       total = length(report_county),
                       denver = sum(report_county %in% 'Denver'),
                       metro = total - denver,
                       denverprop = as.integer(denver / total * 100)
)

actives_print <- merge(x = actives_print,
                       y = actives_rate[ , c("yr_id", "rate_pht")],
                       by = "yr_id",
                       all.x = TRUE)
                       

# Print it with nice names
names(actives_print) <- c("Year", "Total Cases", "Denver", "Other Metro Counties",
                          "% Denver", "Rate per 100k")

dfprintif(actives_print)


```


### TB Cases in Metropolitan Denver, by Quarter

```{r total_cases}


ggplot(actives_qtr, aes(x = plot_qtr, y = freq, group = 1)) +
    geom_point(size = point_size, color = "#31A354") +
    geom_line(size = line_size, color = "#31A354") +
    expand_limits(y = 0) +
    labs(x = "Quarter identified (earliest of date treatment started or date case reported)", 
         y = "Number of cases") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))


```


## Patient Visits by Location
-------------------------------------------



```{r visits_inout}


visits <- query_visits(start_date = first.qtr$start.date,
                       stop_date = last.qtr$end.date)

visits$plot_qtr <- with(visits, paste(visit_yr, " Q", visit_qtr, sep = ""))

locagg <- count(subset(visits, location %in% c("Clinic", "Outreach")),
                vars = c("plot_qtr", "location"))

qtragg <- count(subset(visits, location %in% c("Clinic", "Outreach")),
                vars = "plot_qtr")

qtragg$location <- "All Visits"

visitagg <- rbind(locagg, qtragg)

ggplot(visitagg, aes(x = plot_qtr, y = freq, group = location, color = location)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = location), size = line_size) +
    expand_limits(y = 0) +
    scale_color_manual("Visit Location", values = colorwheel[1:3]) +
    scale_linetype_discrete("Visit Location") +
    labs(x = "Visit date", 
         y = "Number of visits") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))
```


```{r visits_table, results='asis'}

# Cast visits wide
visits_wide <- dcast(visitagg, plot_qtr ~ location, value.var = "freq")

# Pretty-print
names(visits_wide) <- c("Quarter", "Total Visits", "Clinic Visits", "Outreach Visits")

dfprintif(visits_wide)

```


## Screenings for Active and Latent Tuberculosis
-------------------------------------------
```{r screens}

# Query test results for the period
tests <- query_tests(start_date = first.qtr$start.date,
                     stop_date = last.qtr$end.date)

# Convert date to quarter
tests$plot_qtr <- with(tests, paste(test_yr, " Q", test_qtr, sep = ""))

# Aggregate by quarter
testagg <- count(tests, vars = c("test", "plot_qtr"))


# Plot it
ggplot(testagg, aes(x = plot_qtr, y = freq, group = test, color = test)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = test), size = line_size) +
    expand_limits(y = 0) +
    scale_color_manual("Diagnostic", values = colorwheel[1:3]) +
    scale_linetype_discrete("Diagnostic") +
    labs(x = "Diagnostic date", 
         y = "Number of diagnostics") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))
```


```{r screens_table, results='asis'}

# Cast tests wide
tests_wide <- dcast(testagg, plot_qtr ~ test, value.var = "freq")

# Pretty-print
names(tests_wide) <- c("Quarter", "CXRs", "QFTs", "TSTs")

dfprintif(tests_wide)

```





## Treatment Initiation and Completion
-------------------------------------------

```{r plan_start}

# Get the plan starts from the last 9 quarters
plan_starts <- query_tx_plans(start_date = get_quarter(Date = Sys.Date(),
                                                       offset = -9)$start.date,
                              stop_date = get_quarter(Date = Sys.Date(),
                                                      offset = -1)$end.date)


# Create a quarter label
plan_starts$plot_qtr <- with(plan_starts, paste(plan_yr, " Q", plan_qtr, sep = ""))

plan_summary <- ddply(plan_starts, .var = "plot_qtr", .fun = summarise,
                      Active = sum(treat_plan_type %in% "Active"),
                      LTBI = sum(treat_plan_type %in% "LTBI"),
                      RIF = sum(ltbi_drug %in% "RIF"),
                      INH = sum(ltbi_drug %in% "INH"),
                      `INH/Rifapentine` = sum(ltbi_drug %in% "INH/Rifapentine")
)


```





### Treatment Plans Initiated by Type

```{r tx_init}

plan_melt <- melt(plan_summary, id.var = "plot_qtr")

act_ltbi_init <- subset(plan_melt, variable %in% c("Active", "LTBI"))


ggplot(act_ltbi_init, aes(x = plot_qtr, y = value, group = variable, color = variable)) +
    geom_point(size = point_size) +
    geom_line(aes(linetype = variable), size = line_size) +
    expand_limits(y = 0) +
    scale_color_manual("Plan Type", values = colorwheel[1:2]) +
    scale_linetype_manual("Plan Type", values = c(1, 2)) +
    labs(x = "Quarter initiated", 
         y = "Number of patients") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))

```



```{r tx_init_table, results='asis'}

# Pretty names
names(plan_summary) <- c("Quarter", "Active", "LTBI", "Rifampin", "INH", "INH/Rifapentine")

dfprintif(plan_summary)



```


### Completion of Active Therapy

```{r active_comp}


# Query plans again for completion rates, shifting the window back three quarters;
# those nine months allow for active and INH completion.
latest_plans <- query_latest_plans(start_date = get_quarter(Date = Sys.Date(),
                                                           offset = -12)$start.date,
                                   stop_date = get_quarter(Date = Sys.Date(),
                                                           offset = -4)$end.date)


# Create a quarter label
latest_plans$plot_qtr <- with(latest_plans, 
                              paste(plan_yr, " Q", plan_qtr, sep = ""))


# Calculate completion rates for active tx
act_comp <- ddply(subset(latest_plans, treat_plan_type %in% "Active"),
                 .var = "plot_qtr", .fun = summarise,
                 n = sum(completion_status %in% c("Completed", "Not Completed", "Ongoing")),
                 n_complete = sum(completion_status %in% "Completed"),
                 n_incomplete = sum(completion_status %in% "Not Completed"),
                 n_ongoing = sum(completion_status %in% "Ongoing"),
                 per_complete = n_complete / n * 100)


# Pull in the national averages and national targets
nat_avg <- read.csv(file.path("data", "ntip_stats.csv"))

# Add the quarter label
nat_avg$plot_qtr <- with(nat_avg, paste(year, " Q", qtr, sep = ""))

# Subset to only the quarters in the Denver data
nat_avg_sub <- subset(nat_avg, plot_qtr %in% act_comp$plot_qtr)

# Melt it so that average and target are separate records
nat_melt <- melt(nat_avg_sub, 
                 id.var = "plot_qtr", 
                 measure.var = c("national_average", "target_2015"))


# Tidy up the variable labels and column names for stacking
nat_melt$variable <- as.character(nat_melt$variable)
nat_melt$variable[nat_melt$variable %in% "national_average"] <- "National Average*"
nat_melt$variable[nat_melt$variable %in% "target_2015"] <- "National 2015 Target* (93%)"

act_comp$variable <- "All Denver Metro Cases"

names(nat_melt)[names(nat_melt) %in% "value"] <- "per_complete"


# Combine our data, the national average, and the national targets
act_plot <- rbind(act_comp[ , c("plot_qtr", "per_complete", "variable")],
                  nat_melt)

act_plot$per_complete[8] <- 100

# Plot it
ggplot(act_plot, aes(x = plot_qtr, y = per_complete, color = variable, group = variable)) +
 geom_point(size = point_size) +
 geom_line(aes(linetype = variable), size = line_size) +
 expand_limits(y = c(0, 100)) +
 scale_color_manual("Statistic", values = colorwheel[1:3]) +
 scale_linetype_discrete("Statistic") +
 labs(x = "Quarter",
      y = "Percent Completed") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))


```


```{r active_comp_table, results='asis'}

# Set up a version to print
act_print <- act_comp[ , c("plot_qtr", "n", 
                           "n_complete", "n_incomplete", 
                           "n_ongoing", "per_complete")]


# Pretty names
names(act_print) <- c("Quarter", "N Started", 
                      "N Completed", "N Not Completed", 
                      "N Ongoing", "% Complete")


dfprintif(act_print)



```



### Patients on Active Therapy Marked as Incomplete or Ongoing
```{r act_inc,results='asis'}

act_inc <- subset(latest_plans,
                  subset = treat_plan_type %in% "Active" &
                           completion_status %in% c("Ongoing", "Not Completed"),
                  select = c("person_id", "plot_qtr", "treat_plan", 
                             "completion_status", "reason_stopped", "plan_author")
)

dfprintif(arrange(act_inc, plot_qtr, person_id),
          norowmsg = "All active therapy during this period appears to be completed.")

```


# Completion of LTBI Therapy


```{r ltbi_comp}

# Exclude any INH/Rifpentine plans before 2012
plans_trim <- latest_plans[!(latest_plans$ltbi_drug %in% "INH/Rifapentine" & 
                             latest_plans$plan_yr < 2012), ]


# Calculate completion rates for latent tx
ltbi_comp <- ddply(subset(plans_trim, ltbi_drug %in% c("RIF", "INH", "INH/Rifapentine")),
                   .var = c("ltbi_drug", "plot_qtr"),
                   .fun = summarise,
                   
           n = sum(completion_status %in% c("Completed", "Not Completed", "Ongoing")),
           n_complete = sum(completion_status %in% "Completed"),
           n_incomplete = sum(completion_status %in% "Not Completed"),
           n_ongoing = sum(completion_status %in% "Ongoing"),
           per_complete = n_complete / n * 100,
           per_incomplete = n_incomplete / n * 100,
           per_ongoing = n_ongoing / n * 100
)


# Add the national target
ltbi_plot <- rbind(ltbi_comp[ , c("plot_qtr", "ltbi_drug", "n", "per_complete")],
                   data.frame(plot_qtr = sort(unique(ltbi_comp[ , "plot_qtr"])),
                              ltbi_drug = "National 2015 Target* (79%)",
                              n = NA,
                              per_complete = 79)
)


# Plot it
ggplot(ltbi_plot, aes(x = plot_qtr, y = per_complete, group = ltbi_drug, color = ltbi_drug)) +
    geom_point(aes(size = n)) +
    geom_line(aes(linetype = ltbi_drug), size = line_size) +
    expand_limits(y = c(0, 100)) +
    scale_size_continuous("Number of Patients", range = c(3, 7)) +
    scale_color_manual("Regimen", values = colorwheel[1:4]) +
    scale_linetype_discrete("Regimen") +
    labs(x = "Quarter",
         y = "Percent Completed") +
    theme_tb + theme(axis.text.x = element_text(angle = 70, 
                                                hjust = 1.30, 
                                                vjust = 1.3))


```

```{r ltbi_comp_table, results='asis'}

# Set up a version to print
ltbi_print <- ltbi_comp[ , c("ltbi_drug", "plot_qtr", "n", 
                             "n_complete", "per_complete", 
                             "n_incomplete", "per_incomplete",
                             "n_ongoing", "per_ongoing")]


# Pretty names
names(ltbi_print) <- c("LTBI Regimen", "Quarter", "N Started", 
                       "N Completed", "% Complete",
                       "N Not Completed", "% Not Completed",
                       "N Ongoing", "% Ongoing")


dfprintif(ltbi_print)



```



### Patients on LTBI Therapy Marked as Ongoing
```{r ltbi_inc,results='asis'}

ltbi_inc <- subset(latest_plans,
                  subset = treat_plan_type %in% "LTBI" &
                           completion_status %in% "Ongoing",
                  select = c("person_id", "plot_qtr", "treat_plan", 
                             "completion_status", "reason_stopped", "plan_author")
)

dfprintif(arrange(ltbi_inc, plot_qtr, person_id),
          norowmsg = "All LTBI therapy during this period appears to be completed or abandoned.")

```


```{r cleanup}



```
